### 2018-07-18 writing/thesis/pages/abstract.tex
Setup latex template.

\chapter{\abstractname}
Abstract here!

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: