### 2018-11-21 writing/thesis/pages/abstract.tex
Further work on ADER-DG, NS, AMR

\chapter{\abstractname}
This thesis is concerned with the simulation of clouds with the ExaHyPE framework.
ExaHyPE is an engine for hyperbolic PDEs that uses the ADER-DG scheme.

We use the compressible Navier-Stokes (NS) equations with heat conduction to describe the dynamics of clouds.
Due to the diffusive components, the NS-Equations are not hyperbolic.
We thus need to modify parts of the numerical scheme.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: