### 2018-12-16 writing/thesis/chapters/scenarios.tex
Add constants and details to atmospheric scenarios.

\chapter{Scenarios}\label{sec:scenarios}
Unless otherwise noted, we use the constants
\begin{align}
\begin{split}
  \gamma &= 1.4, \\
  \Pr &= 0.7, \\
  c_v &= 1.0
\end{split}
\end{align}
The other constants follow from the definition of \cref{eq:fluid-constants}.

We describe some scenarios in terms of the primitive variables $(\Qrho, \Qv, \pressure)$, these can be converted to the conservative variables by the definition of momentum and the definition of energy.   
\section{A Manufactured Solution}\label{sec:manufactured-solution}
Following~\cite{dumbser2010arbitrary}, the scenario can be described in primitive variables by setting
\begin{align}
\begin{split}
  \pressure(\bm{x}, t) &= \pressure_0 \cos( \bm{k} \bm{x} - \omega t ) + \pressure_b, \\
  \Qrho(\bm{x}, t) &= \Qrho_0 \sin (\bm{k} \bm{x} - \omega t) + \Qrho_b, \\
  \Qv(\bm{x}, t) &= \bm{v_0} \sin(\bm{k} \bm{x} - \omega t).
\end{split}
\end{align}
We set the constants to \( \left(  \pressure_0 = 0.1, \pressure_b = \gamma^{-1}, \Qrho_0 = 0.5, \Qrho_b = 1, 
\bm{v_0} = 0.25 {(1,1)}^\intercal, \bm{k} = \pi/5 {(1,1)}^\intercal \right) \).

We derive a source term by inserting this solution into the \textsc{pde} by using the symbolic math toolkit \textit{Sage}~\cite{sagemath}.
This procedure is called method of manufactured solutions~\cite{salari2000code}.

\section{Classical Scenarios}
In addition to the convergence test, we use standard \textsc{cfd} scenarios to evaluate our simulations.

% \subsection{The first Stokes Problem}
% \todo{Do I actually want to include this? Bit boring scenario}
% \cite{dumbser2010arbitrary}

\subsection{Taylor-Green Vortex}
\begin{align}
  \label{eq:taylor-green}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1\\
  \Qv(\bm{x}, t) &= \exp(-2 \mu t)
  \begin{pmatrix}
    \phantom{-}\sin(x) cos(y) \\
- \cos(x) sin(y) 
    \end{pmatrix} \\
  \pressure(\bm{x}, t) &= \exp(-4 \mu t) \, \frac{1}{4} \left( \cos(2x) + \cos(2y) \right) + C
  \end{split}
\end{align}
\todo{Constants!}
$C = 100/\gamma$ \cite{dumbser2016high}

As initial conditions we use \cref{eq:taylor-green} at time $(t = 0)$ and as boundary conditions we impose the solution and the corresponding gradients.
Note that this solution is not exactly correct, as it assumes an incompressible flow.
Due to the low mach number a compressible flow approaches this solution.

\subsection{Lid-Driven Cavity Flow}
The lid-driven cavity flow is a very simple testing scenario for the Navier-Stokes equations.
We use a constant pressure of (what exactly) and a constant density $(\Qrho = 1.0)$ for the entire domain.
The flow is then driven entirely by the boundary conditions.
They consist in four no-slip walls where the upper wall has a constant velocity $\Qv = (1.0, 0)$.
This scenario is surprisingly difficult for an \textsc{ader-dg} scheme as the boundary conditions lead to a discontinuity at the top left corner.
There, the velocity of the wall is piece-wise constant.
Therefore, this scenario is a good choice to test the stability of the finite volume limiter.

\subsection{ABC-Flow}
We use the Arnold–Beltrami–Childress (\textsc{abc}) flow as a simple example of a three-dimensional scenario~\cite{tavelli2016staggered}.
This scenario has, unlike the three-dimensional version of the Taylor-Green vortex, at least for the incompressible case an analytical solution
\begin{align}
  \label{eq:abc-flow}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1\\
  \Qv(\bm{x}, t) &= \phantom{-} \exp(-1\mu t)
  \begin{pmatrix}
    \sin(z) + \cos(y)\\
    \sin(x) + \cos(z)\\
    \sin(y) + \cos(x)
  \end{pmatrix} \\
  \pressure(\bm{x}, t) &= -\exp(-2 \mu t) \, \left(\cos(x)\sin(y) + \sin(x)\cos(z) + \sin(z)\cos(y)\right)
  + C
  \end{split}
\end{align}
Thus, we can simulate this without periodic boundary conditions.
\todo{Analytical solution!}

\subsection{Radialsymmetric CJ Detonation Wave}
We investigate the radialsymmetric CJ detonation wave scenario from~\cite{helzel2000modified}.
This scenario for the reactive Euler-equations is set up in a 2D box of size $[-0.5, 0.5] \times [-0.5, 0.5]$.
Inside a circle of radius $0.3$ we have totally unburnt gas, outside of this circle we have burned gas.
Let $\alpha =  \operatorname{atan2}(y,x)$ be the angle in polar coordinates.
We use the initial conditions
\begin{align}
  \begin{split}
  \Qrho_u &= 0.887565\\
  \Qv_u &= -0.577350
  \begin{pmatrix}
     \cos (\alpha)\\
     \sin (\alpha)
  \end{pmatrix} \\
  \pressure_u &= 0.191709 \\
  \QZZ_u &= 1
  \end{split}
\intertext{for the burnt gas and}
\begin{split}
  \Qrho_b &= 1.4\\
  \Qv_b &= 0\\
  \pressure_b &= 1 \\
  \QZZ_b &= 0
  \end{split}
\end{align}
for the unburnt gas.

Similar to~\cite{helzel2000modified}, we use the constants
\begin{align}
  \begin{split}
  \gamma &= 1.4, \\
  \Pr &= 0.7, \\
  c_v &= 2.5, \\
  c_p &= 3.5, \\
  R &= 1.0.
  \end{split}
\end{align}

\section{Atmospheric Scenarios}
The scenario is described in terms of potential temperature $\potT$
\begin{equation}
  \potT = T \frac{p_0}{p}^{R/c_p},
\end{equation}
where $p_o = \SI{10e5}{\Pa}$ is the reference pressure.
Solving for $T$ leads to
\begin{equation}
  \label{eq:potTToT}
  T = \potT (\frac{p}{p_0})^{R/c_p}.
\end{equation}
To allow for an easier description of the initial conditions we split the $\potT$ into a background state $\backgroundPotT$ and a pertubation $\pertubationPotT$
\begin{equation}
  \label{eq:potT-split}
  \potT = \backgroundPotT + \pertubationPotT.
\end{equation}

The initial conditions for pressure and density can be computed directly from the \textsc{pde} once few assumptions have been given.
We assume that $\backgroundPotT$ is constant over the entire domain.
Inserting the definition of potential temperature, given by \cref{eq:potTToT}, into the equation of state (see \cref{eq:eos}) leads us to
\begin{equation}
  \pressure(z) = \Qrho(z) R {\left( \backgroundPotT \frac{p}{p_0}\right)}^{R/c_p},
\end{equation}
which we solve for $p(z)$.
After some algebra and simplifying, we arrive at
\begin{equation}
 p(z) = p_0 \left( \frac{R \backgroundPotT \rho(z)}{p_0} \right)^{\gamma},
\end{equation}
or equivalently
\begin{equation}
  \label{eq:eos-potT}
\Qrho(z) = \frac{p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}.
\end{equation}
We want to consider flow in hydrostatic equilibrium, i.e.\ flow where the force originating from the pressure-gradient is exactly balanced by gravitation.
Inserting this assumption in the momentum equation in $z$-direction and using \cref{eq:eos} once again, we arrive at the ordinary differential equation (\textsc{ode})
\begin{align}
  \label{eq:hydrostatic-balance-potT}
  \begin{split}
  \frac{d}{d z} p{\left (z \right )} &= -g \Qrho(z)
                                     = 
        - \frac{g p_{0}^{\frac{R}{c_{p}}} p^{\frac{1}{\gamma}}{\left (z \right )}}{R \backgroundPotT}\\
  p{\left (0 \right )} &= p_0
  \end{split}
\end{align}
where we use the reference pressure $p_0$ as the pressure on ground level.
This \textsc{ode} can be now simply solved by separation of variables.
We then obtain after simplifying the solution
\begin{align}
  \label{eq:hydrostatic-background-pressure}
p(z) &= \left(\left(1 - \frac{1}{\gamma}\right) \left(C - \frac{g p_{0}^{\frac{R}{c_{p}}} z}{R \backgroundPotT}\right)\right)^{\frac{c_{p}}{R}},\\
\shortintertext{with constant of integration}
 C &= \frac{c_{p} p_{0}^{\frac{R}{c_{p}}}}{R}.
\end{align}

We can now compute the initial conditions for a scenario where the background potential energy is in hydrostatic equilbrium with a pertubation that is not in equilbrium.
We first compute the background pressure with background potential temperature $\backgroundPotT = \text{const.}$ using \cref{eq:hydrostatic-background-pressure}.
Next, we compute the pertubated potential temperature $\potT$ and use \cref{eq:potTToT} to convert it to temperature.
The density $\Qrho(\bm{x})$ can now be evaluated with \cref{eq:temperature}.
Finally, we compute the energy by inserting pressure and density into the equation of state~(\ref{eq:eos}).

This leaves us with one problem:
The adiabatic no-slip boundary conditions defined by~\vref{eq:no-slip} are no longer valid.
Concretely, we need to impose a viscous heat flux to ensure that the atmosphere stays in hydrostatic balance~\cite{giraldo2008study}
\todo{Flux correct? Fix notation}
\begin{equation}
  \label{eq:atmosphere-bc}
  F^\text{visc} = \kappa \pdv{\overline{T}}{z} =
\frac{R \overline{\theta} \left(\frac{\overline{p}{\left (z \right )}}{p_{0}}\right)^{\frac{R}{c_{p}}} \frac{d}{d z} \overline{p}{\left (z \right )}}{c_{p} \overline{p}{\left (z \right )}}.
\end{equation}
The background pressure can be reconstructed by \cref{eq:hydrostatic-background-pressure}, its derivative is given by \cref{eq:hydrostatic-balance-potT}.

We use the following set of fluid dependent constants for these scenarios
\begin{align}
  \begin{split}
    \gamma &= 1, \\
    \Pr &=  0.71, \\
    R &= 287.058, \\
    p_0 &= 10000,
  \end{split}
\end{align}
the other constants are evaluated following \cref{eq:fluid-constants}.

\subsection{Two Bubbles}
As a simple test case we consider a scenario where we have one large warm air bubble in the middle, and a small cold bubble on the top.
This scenario was first published by \citeauthor{robert1993bubble}~\cite{robert1993bubble}.
We use a domain of $\SI{1}{\km} \times \SI{1}{\km}$.
The initial conditions of the scenario are given by a pertubation of a hydrostatic equlibrium.
We use a background potential temperature of $\backgroundPotT = 300$.

Let
\begin{equation}
  \label{eq:radius}
  r^2 = (x - x_0)^2 + (z - z_0)^2
\end{equation}
denote the difference between spatial positions $x,z$ and the center of a bubble $x_c, z_c$.
The pertubation is given by the function
\begin{equation}
  \label{eq:bubbles-pertubation}
  \pertubationPotT =
  \begin{cases}
    A & r \leq a \\
    A \exp \left( - (r-a)^2/s^2 \right) & r > a,
    \end{cases}
\end{equation}
where $A$ denotes the maximal pertubation, $a$ is the size of the bubble and $s$ is the decay rate.
We have two bubbles, with the values
\begin{equation}
  \label{eq:bubbles-values}
\begin{alignedat}{6}
  & \text{warm:} \qquad && A = 0.5 \quad&& a = \SI{150}{\m} \quad&& s = \SI{50}{\m} \quad&& x_c = \SI{500}{\m} \quad&& z_c = \SI{300}{\m},\\
  & \text{cold:} \qquad && A = -0.15 \quad&& a = \SI{0}{\m} \quad&& s = \SI{50}{\m} \quad&& x_c = \SI{560}{\m} \quad&& z_c = \SI{640}{\m}.
  \end{alignedat}
\end{equation}

Similar to~\cite{muller2010adaptive}, we use a constant viscosity of $\mu = 0.1$ to regularize the solution.

\subsection{Density current}
A similar scenario is the density current~\cite{straka1993numerical} by \citeauthor{straka1993numerical}.
In this scenario we only have one cold bubble.
We use the same procedure as before to compute the initial conditions, again with $\backgroundPotT = 300$ but with a pertubation of the form
\begin{align}
  \label{eq:cos-pertubation}
  \pertubationPotT &= \begin{cases}
    \nicefrac{A}{2} \left[ 1 + \cos(\pi r) \right] & r \leq a \\
    0 & r > a
    \end{cases}
\intertext{with}        
        r^2 &=  \left( \frac{x - x_c}{x_r} \right)^2 + \left( \frac{z -z_c}{z_r} \right)^2
\end{align}
and constants
\begin{equation}\label{eq:density-bubble}
  A = -15 \quad a = \SI{1}{\m} \quad x_r = \SI{4000}{\m} \quad z_r = \SI{2000}{\m} \quad x_c = \SI{0}{\m} \quad z_c = \SI{3000}{\m}.
\end{equation}
Similar to~\cite{giraldo2008study} we use a constant viscosity of $\mu = 75$.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: