### 2019-01-11 writing/thesis/pages/abstract.tex

\chapter{\abstractname}
This thesis is concerned with the simulation of clouds with the ExaHyPE framework.
ExaHyPE is an engine for hyperbolic PDEs that uses the ADER-DG scheme.

We use the compressible Navier-Stokes (NS) equations with heat conduction to simulate the dynamics of clouds.
Due to the diffusive components, the NS-Equations are not hyperbolic.
We thus need to modify parts of the numerical scheme.

Furthermore, we extend the equation set to the reactive compressible Navier-Stokes equations which include a source term that models chemical reactions.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: