### 2019-01-20 writing/thesis/pages/abstract.tex

\chapter{\abstractname}
This thesis is concerned with the simulation of clouds with the ExaHyPE framework.
ExaHyPE is an engine for hyperbolic partial differential equations that uses the ADER-DG scheme.

We use the compressible Navier-Stokes equations for computational fluid dynamics.
Due to the diffusive components, the Navier-Stokes equations are not hyperbolic.
We thus need to modify parts of the numerical scheme.
Furthermore, we extend the equation set to the reactive compressible Navier-Stokes equations which include a source term that models chemical reactions.

We develop a gradient-based, global adaptive mesh refinement indicator that finds cells with unusually steep gradients.

Our implementation is evaluated for standard two and three dimensional fluid dynamics test cases, for flows over a realistic background atmosphere and for a reactive detonation wave.
Furthermore, we perform a convergence test and a time-to-solution benchmark that evaluates our proposed AMR criterion.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: