### 2019-01-20 writing/thesis/pages/acknowledgements.tex

\thispagestyle{empty}

% \vspace*{20mm}

% \begin{center}
% %{\usekomafont{section} Acknowledgments}
% \end{center}

% \vspace{10mm}

% \renewcommand{\dictumwidth}{\textwidth}
% \setchapterpreamble{\dictum[Booktitle]{
%     Quote!
% }}
% %\addcontentsline{toc}{chapter}{Acknowledgments}
\chapter{Acknowledgments}
\vspace{1em}

Acknowledgements here.
\cleardoublepage{}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: