### 2019-01-21 writing/thesis/pages/acknowledgements.tex

\thispagestyle{empty}

% \vspace*{20mm}

% \begin{center}
% %{\usekomafont{section} Acknowledgments}
% \end{center}

% \vspace{10mm}

% \renewcommand{\dictumwidth}{\textwidth}
% \setchapterpreamble{\dictum[Booktitle]{
%     Quote!
% }}
% %\addcontentsline{toc}{chapter}{Acknowledgments}
\chapter{Acknowledgments}
\vspace{1em}
I want to thank Prof.\ Bader for the chance to work on this exciting topic.
I also want to thank my advisor Leonhard Rannabauer, whose guidance made this thesis not only possible but also enjoyable.

My gratitude extends to my friends and family, who supported me during the entire thesis.
Special thanks go to my friend William who proof-read the thesis.
\cleardoublepage{}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: