### 2019-01-21 writing/thesis/pages/acknowledgements.tex

\thispagestyle{empty}

% \vspace*{20mm}

% \begin{center}
% %{\usekomafont{section} Acknowledgments}
% \end{center}

% \vspace{10mm}

% \renewcommand{\dictumwidth}{\textwidth}
% \setchapterpreamble{\dictum[Booktitle]{
%     Quote!
% }}
% %\addcontentsline{toc}{chapter}{Acknowledgments}
\chapter{Acknowledgments}
\vspace{1em}
I want to thank Prof.\ Michael Bader for the chance to work on this exciting topic.
I am particularly grateful to my advisor Leonhard Rannabauer, whose guidance made this thesis not only possible but also enjoyable.
Additionally, I wish to acknowledge the support of Dominic Charrier, who was of great help with the implementation in \exahype{}.

My gratitude extends to my friends and family, who supported me during the entire thesis.
Special thanks go to my friend William who proofread this document.
\cleardoublepage{}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../main"
%%% End: